---
title: "Analysis of ColdPlay band musics"
subtitle: "My subtitle if needed"
author: 
  - Aakash Vaithyanathan
  - Kevin Cai
thanks: "Code and data are available at: [Code](https://github.com/aakash2002/STA304-W5)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(here)

data <- readRDS(here::here("data/analysis_data/coldplay_cleaned_data.rds"))
```

# Introduction[intro]


# Data & Results[data]

## Energy of song and it's duration
```{r}
#| label: fig-energy-duration
#| fig-cap: Relationship between energy of song and its duration in minutes
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = duration_min, y = energy)) +
  geom_point() +
  labs(title = "Relationship between key notes played and energy of song for 2000-2024 years",
       x = "Duration in Minutes", y = "Energy of Song") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))


```
The above graph shows the relationship between the duration of a song to its energy. The energy of a song corresponds to its combined intensity and it's vigor (@energyOfSong). As we can see, many songs have high energy values between the range 3 to 5 minutes. The band Coldplay (@coldplay) is an alternative rock style band and the findings make sense as these songs have a strong impact on our relative mood and the energy we experience. We also notice that as the duration of the song decreases, many albums/songs have relatively low energy levels.


\newpage


# References[ref]

# Appendix


